---
title: "RRR3 Intro to Knitr; IRRR1 Knitr"
author: "Jessica Minnier, BSR Reproducible Research Journal Club"
date: "August 17, 2016"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [RRRR3](https://englianhu.files.wordpress.com/2016/01/reproducible-research-with-r-and-studio-2nd-edition.pdf). Getting Started with R, Rstudio, and **knitr**

- Last week:
    + 3.1 -- using R: the basics
    + 3.2 -- using RStudio (questions?)
- This week:
    + 3.3 -- using *knitr* and *r markdown*: the basics

## [IRRR1](https://books.google.com/books?id=WVTSBQAAQBAJ&pg=PA3&source=gbs_toc_r&cad=4#v=onepage&q&f=false) knitr: A comprehensive tool for reproducible research in R – Yihui Xie

- A short version of Yihui’s book *Dynamic Documents with R and Knitr*
- [Knitr website](http://yihui.name/knitr/)
- [FAQ's here](https://github.com/yihui/knitr/blob/master/FAQ.md)
- [Watch his 5 minute intro video](http://www.screenr.com/qcv8)
- From a computer science perspective
- Knitr package in R
    + built upon previous tools (WEB, Noweb, Sweave)
    + framework redesigned to allow for easy and fine control of many aspects of a report
    + Sweave is a subset of knitr in terms of features
- Main design philosophy: “make reproducible research easier and more enjoyable than the common practice of cut-and-paste results”

## Knitr – R Markdown (Rmd)

- Rmd is one of the documents that knitr supports (and simplest).
- If we have a .Rmd file with code and text, we can use the `knit` command to run the R code and convert the document to html:
```
  library(knitr)
  knit("myfile.Rmd")
```
- or we can use the "Knit HTML" button in Rstudio which uses the `knitr` package to do the same thing.

## Behind the scenes of Knitr (IRRR1)

Three parts:

- parser: finds the code in your document based on special patterns of text (such as code that starts with ```{r} and ends with ``` )
- evaluator: evaluates the R code (or other type of code), returns a list of output (including errors, messages, normal output)
    + Note we can turn some of this off using **code chunk options**
- renderer: renders the text, code, output (or a subset of those if we choose) as .md+.html code (or .docx, or .tex+.pdf)

## Code Chunks

This is a code chunk with no name, no options, but it will run:

    ```{r}`r ''`
    x = rnorm(100)
    plot(x)
    ```

This is a code chunk with name and chunk options:

    ```{r normaloutput, fig.width=5, echo=FALSE}`r ''`
    x = rnorm(100)
    plot(x)
    ```

## Features of Knitr (IRRR1)

- code highlighting
- graphics rendering
    + figure files also can be saved as separate files (chunk option `fig.keep='all'`)
    + animations: `fig.show='animate'`
    + figure specifications: `fig.width` and `fig.height` for example
- cache: `cache=TRUE`
    + for slow code, can save the output objects from a chunk of code in the "cache"
    + directly loads results from a previous run
    + if you change the code in a cached chunk, the code is rerun and resaved
    + be careful with caching, don't load packages or set knitr options, see details on [knitr website](http://yihui.name/knitr/demo/cache/)
- chunk hooks (see 1.4.1)
- other languages supported (python, C++, bash, etc)


## YAML

```
---
title: "Habits"
author: "Jane Doe"
output:
  html_document:
    theme: united
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
---
```

- The header of your R Markdown document has the title options
- See [description of the many options](http://rmarkdown.rstudio.com/html_document_format.html)
- Add table of contents `toc`, change theme appearance `theme` or with custom CSS files,
include other code in header (i.e. latex code or html code), figure options with `fig_width` and `fig_height`
- Can change some with "Output options" button next to "Knit HTML" button in Rstudio

## Code chunk options

- code options are like function options (i.e. `eval=FALSE`)
- if you've defined a TRUE/FALSE variable called `dothis` earlier in the document, 
you can even use `eval=dothis`
- See [description of the many, *many* options](http://yihui.name/knitr/options/)
- must be one line (no line breaks)
- avoid spaces `' '` and periods `.` in chunk labels and directory names

An example:

    ```{r eval=TRUE, fig.width=5, echo=FALSE} `r ''`
    x = rnorm(100)
    plot(x)
    ```
Also set global cunk options in the beginning

    ```{r setup, include=FALSE} `r ''`
    knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
    ```

## [Code chunk options examples](http://yihui.name/knitr/options/#chunk_options)

Common:
```
echo=FALSE        Don't include the code
results="hide"    Don't include the output 
include=FALSE     Don't show code or output
eval=FALSE        Don't evaluate the code at all
warning=FALSE     Don't show R warnings
message=FALSE     Don't show R messages 
fig.width=#       Width of figure
fig.height=#      Height of figure
fig.path="Figs/"  Path for figure files
```

Advanced:
```
cache=TRUE        Cache code chunk
results="asis"    Use with pandoc.table or other html generating code
tidy=TRUE         Tidy up your code output
fig.show="hide"   Generate plot files but hide them in report
dev="png"         Change file type of figure/plot
```

## Extract R code (`purl()`)

You can extract all R code from your knitr document into an R file, with or without documentation.

```{r purlthis, eval=FALSE}
library(knitr)
setwd("slides-info-reproducible-research/study-group-2016/2016_08_10_intro-rstudio-knitr/")
purl("minnier-2016-08-17-knitr.Rmd", output="minnier-2016-08-17-knitr-no-doc.R")
purl("minnier-2016-08-17-knitr.Rmd",
     output="minnier-2016-08-17-knitr-with-doc.R",
     documentation = 2)
```

(Since all the code in this file has `eval=FALSE` the R code is commented out. Try with a regular .Rmd file to see what evaluated code looks like.)

## R Notebooks (`spin()`)

- .R file with pure code (no chunks) but with specially formatted comments that are then
parsed and compiled inito a report similar to knitr.

```
#' # This is a header
#' 
#' I can write comments like this that will be turned into
#' markdown text.

#+ chunkname, fig.width=10
x = rnorm(100)
# regular R comment, within my code chunk
plot(x)

#' This is the end of the report. Now click "Compile Notebook"
#' button in Rstudio. Or use `spin()` function
#' in knitr package, and we will have a weaved html document! 
#' Not as pretty as regular .Rmd+knitr but we can add
#' many options to make it just as nice, even YAML!
#'
```

You can compile this in Rstudio (notebook button) or R with the function `spin()`:
```{r eval=FALSE, echo=TRUE}
library(knitr)
spin("example_notebook_code.R")
```

## R and `knitr`

Instead of using Rstudio's built in buttons, we can perform all the same tasks in pure
R code.

To knit a .Rmd file to .md and then convert to .html we can use the code:
```{r eval=FALSE, echo=TRUE}
setwd("~/Documents/") # change working directory to directory with .Rmd file
knit(input = "Example.Rmd", output = "Example.md") # .Rmd to .md
mardownToHTML(file = "Example.md", output = "Example.html") #.md to .html
```

or, in one line:
```{r eval=FALSE, echo=TRUE}
knit2html(file = "~/Documents/Example.Rmd", output = "~/Documents/Example.html")
```

## R and `rmarkdown`

If we have a markdown file with YAML heading specifying the type of output and options, we can also use
the `rmarkdown` package in R:

```{r eval=FALSE, echo=TRUE}
render(file = "~/Documents/Example.Rmd")
```

we can even render multiple types of files from the same document:
```{r eval=FALSE, echo=TRUE}
render("Example.Rmd", c("pdf_document", "html_document"))
```


## Now let's see an example!

0. [Install git](https://git-scm.com/downloads)
1. File --> New Project --> Version Control --> Git
2. Repository URL:
https://github.com/jminnier/DataSharingPolicies.git
3. Project Directory Name: `WhateverYouWant`
4. Create project as subdirectory of: `SomewhereOnYourComputer/`
5. X Open in new session
6. **Create Project**
7. Open `analysis-code/analysis-datasharing.Rmd`
8. Knit away!
9. Also try running `render-analysis-report.R`

**What does this code do?**
```{r library, eval=FALSE}
source("render-analysis-report.R")

setwd("analysis-code")
purl("analysis-datasharing.Rmd", output = "mycode.R")
purl("analysis-datasharing.Rmd", output = "mycode-withdoc.R", documentation = 2)

knitr::knit2html("analysis-data-sharing.Rmd", output = "myoutput.html")

rmarkdown::render("analysis-datasharing.Rmd",
                  output_format = c("html_document", "pdf_document"),
                  params = list(showcode=TRUE))
```


## Rmd Reference Guides


- [Longer Reference Guide](http://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
- [Cheatsheet](http://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)



