---
title: "RRRR 8. Statistical Modeling and Knitr"
author: "Jeong Lim"
date: "August 31, 2016"
output: html_document
---


## Introduction 
Focus on how to make your analysis really reproducible

- Incorporating analyses into the Markup : `chunk` options
- Computationally intensive code: chunk option `cache`
- Including non-R code in code chunks (ex. SAS code)
    
    
## Code Chunks
 
 
    ```{r}`r ''`
    
    ```

 Code chunk with option

    ```{r NAME, OPTIONS HERE}`r ''`
    
    ```
    
 Global option

    ```{r,  include=FALSE }`r ''`
    knitr::opts_chunk$set(OPTIONS HERE)
    ```

## [Chunk options](http://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)


```
echo=FALSE        Don't include the code
results="hide"    Don't include the output 
include=FALSE     Don't show code or output
eval=FALSE        Don't evaluate the code at all

collapse=TRUE     Collapse all the source and ouput blocks into a single block    
warning=FALSE     Don't show R warnings
message=FALSE     Don't show R messages 
error=FALSE       Don't show R error 
cache=TRUE        Cache code chunk
tidy=TRUE         Reformat code in a tidy way
comment=NA        Remove ## 

```


## Chunk options example 
- numeric vector to select source code ;  e.g) `echo=1`, `echo=c(1,3)`, `echo=-(1:2)`

code chunk without option
      
    ```{r withoutoption}`r ''`
    cars[1:4,]
    sum(cars$speed)
    sum(cars$dist)
    mean(cars)
    plot(cars$speed, cars$dist)
    ``` 
It looks like 

```{r withoutoption}
cars[1:4,]
sum(cars$speed)
sum(cars$dist)
mean(cars)
plot(cars$speed, cars$dist)
```
  
code chunk with options
  
    ```{r withoption, echo=2:3,  warning=FALSE, collapse=TRUE, comment=NA, fig.align='center', fig.width=4} `r ''`
    cars[1:4,]
    sum(cars$speed)
    sum(cars$dist)
    mean(cars)
    plot(cars$speed, cars$dist)
    ```

It looks like 
  
```{r withoption, echo=2:3,  warning=FALSE, collapse=TRUE, comment=NA, fig.align='center', fig.width=4} 
cars[1:4,]
  sum(cars$speed)
sum(cars$dist)
mean(cars)
plot(cars$speed, cars$dist)
```



## [Chunk options example:` tidy`](http://yihui.name/formatR/)

- `tidy_source()` function in the formatR package make code tidy to improve readability 

    +  long lines of code and comments are reorganized into appropriately shorter ones
    +  spaces and indent are added where necessary
    +  comments are preserved in most cases
    +  the number of spaces to indent the code (i.e. tab width) can be specified (default is 4)
    +  an else statement in a separate line without the leading } will be moved one line back
    +  = as an assignment operator can be replaced with <-
    +  the left brace { can be moved to a new line



```{r}
library(formatR)
source("C:/Users/limje/Desktop/Reproducible/Jeong/UglyScript.R")
tidy_source("UglyScript.R", file="BeautifulScript.R")
```




## Showing code & result inline

-  Inline static code : single backticks
         
  ex) This is example R code : ` MeanRiver <- mean(rivers) `
  
     
-  Inline dynamic code : r after the first single backtick 


  ex) The mean length of 141 major rivers in North America is `r round(mean(rivers), digits=0)`
     
     

## Dynamically including Modular Analysis files
-  Long R source code chunks
-  Same code in muliple documents (eg.article and slide show presentation)
-  Interested in specific parts of analysis

Source from a local file

```{r, eval=FALSE}
source("C:/Users/limje/Desktop/Reproducible/Jeong/MainAnalysis.R")
```


Source from a secure URL

```{r,fig.width=4, fig.height=4}
library(devtools)
source_url("http://bit.ly/1D5p1w6")
```

SHA-1 hash is a unique number for the file. If the file changes, its SHA-1 hash will change



## Computationally intensive analysis: `cache=TRUEE`
 * knitr will first check if there's previously cached results
 * `cache.path` : set the cache directory
 * Only be run when the chunk's contents change or chunk option change
 
 
create an object Sample to a file called Sample.RData

       ```{r gen-data, cache=TRUE} `r ''`
       # create data
       Sample<-rnorm(n=1000, mean=5, sd=2)
       # save sample
       save(Sample, file="Sample.RData")
       ```

latter code chunk for creating the histogram 

       ```{r histgram, cache=TRUE, dependson='gen-data'}`r ''`
       # load Sample
       load(file="Sample.RData")
       # create histogram
       hist(Sample)
       ```

## Running SAS code using chunk option `engine`

      ```{r, engine="sas", engine.path="C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"} `r ''`
      proc means data=sashelp.class;
      run;
      ```
SAS code and output 

```{r, engine="sas", engine.path="C:/Program Files/SASHome/SASFoundation/9.4/sas.exe", comment=""}
proc means data=sashelp.class;
run;

```


SAS HTML output (only works with HTML )

```{r, engine="sas", engine.path="C:/Program Files/SASHome/SASFoundation/9.4/sas.exe", echo=FALSE, results='hide'}
ods noproctitle;
ods graphics;
ods html file="sasoutput.html" (no_top_matter no_bottom_matter) style=journal;
proc means data=sashelp.class;
run;
```

```{r, child="sasoutput.html", echo=FALSE}
```




## Reproducibly Random : 'set.seed'
```{r, fig.width=4, fig.height=4}
set.seed(123)
Draw1<-rnorm(1000, mean=0, sd=2)
summary(Draw1)
hist(Draw1)

set.seed(125)
Draw2<-rnorm(1000, mean=0, sd=2)
summary(Draw2)
hist(Draw2)
```