---
title: "RRRR 9. Showing Results with Tables"
author: "Solange Mongoue"
date: "September 7, 2016"
output: slidy_presentation
---


## Outline 
Learn how to dynamically connect statistical results with tables in your presentation documents written in LaTeX and Markdown 

Focus on three tools in this chapter:  

- `kable` function from *knitr*   
  creates documents from data frames for both LaTeX and Markdown/HTML documents  
- the *xtable* package  
  does the same but much more customizable  
- the *texreg* package  
  procuces publication-quality tables. can combine results from multiple models into one table  
    
    
## Basic knitr Syntax for Tables

`results` option is the most important chunk option  

4 values  

- `hide` hides results of your code chunk  
- `asis` includes table created from R objects. writes raw markup form of the table into the presentation document  
- `markup` uses output hook to mark up the results in a predefined way  
- `hold` collects all the output and prints at the end of the chunk  



## Tables in LaTeX

### The center environment

```
\begin{center}

  This is a center environment.    
  
\end{center}   
```

### The tabular environment

To begin, type `\begin{tabular}{TABLE_SPEC}`.  

`{TABLE_SPEC}` allows you to specify the number of columns and the text alignment in each column.  

Example: to create a table with three columns, the first is left-justified and the latter two are center-justified, type

```
\begin{tabular}{l | c c}

```

- If we want to add a vertical line between two columns, we can place a vertical bar `|` between the columns arguments  
- Columns are delimited with `&`  
- Use two backslashes to start a new line `\\`  
- To add horizontal line, use `\hline`  
- Close the tabular environment with `end{tabular}`

Full laTeX code

```
\begin{tabular}{l | c c}
  \hline
  Observation & Variable1 & Variable2 \\
  \hline \hline
  Subject1 & a & b \\
  Subject2 & c & d \\
  Subject3 & e & f \\
  Subject4 & g & h \\
  \hline
\end{tabular}

```



## Tables in LaTeX

### The table float environment
 
 Allows us to separate a table from the text, specify its location and give it a caption.  
 
- To begin use `\begin{table}[POSITION_SPEC]`
- The `POSITION_SPEC` argument allows us to determine the location of the table  
    * `h` for here, i.e. where the table is written in the text  
    * `t` for top of the page  
    * `b` for bottom of the page  
- Use the `\caption` command to set a title for the table  
- Use `\label` command to declare a cross-reference key for the table. To reference the table in the test, simply type `\ref{KEY}`, where `\KEY` is set with the `\label` command  
- A table environment is closed with `\end{table}`
 
LaTeX code
 
```
\begin{table}[t]
    \caption{Example Simple LaTeX Table}
    \label{ExLaTeXTable}
    \begin{center}
      \begin{tabular}{l | c c}
        \hline
        Observation & Variable1 & Variable2 \\
        \hline \hline
        Subject1 & a & b \\
        Subject2 & c & d \\
        Subject3 & e & f \\
        Subject4 & g & h \\
        \hline
      \end{tabular}
    \end{center}
\end{table}

```


## Tables in Markdown/HTML

### Markdown tables

- Much simpler than laTeX's syntax  
- No need to declare a new environment, just start typing  
- Columns are delimited with vertical bars (|). Not necessary to line up the vertical bars  
- Rows are started with a new line  
- Specify each column's text justification using colons on the dashed row  

Example: this code will create the left-center-center justified formatted table

``` {r, echo=TRUE, eval=FALSE}
### Example Simple Markdown Table
Observation | Variable1 | Variable2
----------- | --------- | ---------
Subject1    | a         | b
Subject2    | c         | d
Subject3    | e         | f
Subject4    | g         | c
```

### Example Simple Markdown Table
Observation | Variable1 | Variable2
---------- | ------- | -------
Subject1    | a         | b
Subject2    | c         | d
Subject3    | e         | f
Subject4    | g         | c

## Tables in Markdown/HTML

### HTML tables

- `textreg` function creates table with HTML syntax 
- HTML uses **tables** element 'tags' to begin (`<table>`) and end tables (`</table>`) 
- Not required to tab the content of a table element 
- Specify element attributes inside start tags i.e. `<table, border="1">` to add a border to a table 
- Table rows are put inside of `tr` (table rows) element tags 
- Individual cells are delimited with `td` (standard cell) tags 
- Delimit a table's header row from its body with the `thead` and `tbody` tags 
- Table captions can be included with `caption` tags

Example: Recreate the same left-center-center justified formatted table

```
<table>
  <thead>
    <tr>
        <td>Observation</td> <td>Variable1</td> <td>Variable2</td>
    </tr>
  </thead>
  <tbody>
    <tr>
        <td>Subject1</td> <td>a</td> <td>b</td>
    </tr>
    <tr>
        <td>Subject2</td> <td>c</td> <td>d</td>
    </tr>
    <tr>
        <td>Subject3</td> <td>e</td> <td>e</td>
    </tr>
    <tr>
        <td>Subject4</td> <td>f</td> <td>f</td>
    </tr>
  </tbody>
</table>
```

<table>
<thead>
<tr>
<td>Observation</td> <td>Variable1</td> <td>Variable2</td>
</tr>
</thead>
<tbody> <tr>
<td>Subject1</td> <td>a</td> <td>b</td>
</tr>
<tr>
<td>Subject2</td> <td>c</td> <td>d</td>
</tr> <tr>
<td>Subject3</td> <td>e</td> <td>e</td>
</tr>
<tr>
<td>Subject4</td> <td>f</td> <td>f</td>
</tr>
</tbody>
</table>



## Creating Tables from Supported Class R Objects

Use `kable` funtion, __xtable__ and __texreg__ packages. 

They work easily with specific object classess that their designers explicitly support.

### `kable` for Markdown and LaTeX

Converts matrices and data frames into tables for Markdown, HTML and LaTeX
```{r include=FALSE}
library(knitr)
```
```{r , results='asis', echo=TRUE}
kable_ex <- data.frame(
            Observation = c("Subject1", "Subject2",
                            "Subject3", "Subject4"),
            Variable1 = c("a", "c", "e", "g"),
            Variable2 = c("b", "d", "f", "c")
)
kable(kable_ex, caption = "Example kable Table", format = "markdown")
```

- Can specify new column and row names by passing character vectors to `col.names` and `row.names` 
- Use `digits` argument to specify the number of digits after the decimal place 
- Use `format='latex'` to create a LaTeX formatted table 

## Creating Tables from Supported Class R Objects

### `xtable` for LaTeX and HTML 

- *kable* has limited customizability 
- **xtable** package can create more customizable tables from a wider variety of R objects, including statistical model objects 

Example: `lm` (linear model) command creates model summaries of the `lm` class. Let's create a simple linear regression using the swiss data frame and `lm` command.

```{r, echo=TRUE, comments='##'}
# Fit simple linear regression model
M1 <- lm(Examination ~ Education, data = swiss)
# Show M1 class
class(M1)
# Show summary of M1 model object
summary(M1)
```
  
- To find a full list of object classes that *xtable* supports, type `methods(xtable)' after loading the package  

```{r, results='asis', echo=TRUE}
library(xtable)
methods(xtable)
```

## Creating Tables from Supported Class R Objects

#### `xtable` for LaTeX

Let's create a table summarizing the estimates from the M1 model obsject. 
The code below will generate the LaTeX syntax.

```{r, echo=TRUE}
# Load xtable
library(xtable)

# Create LaTeX table from M1 and show the output markup
xtable(M1, caption = "Linear Regression,
        Dependent Variable: Exam Score",
        label = "BasicXtableSummary",
        digits = 1)

```

## Creating Tables from Supported Class R Objects

#### `xtable` for Markdown/HTML

- Use *xtable* and the `print.xtable` command to create tables for Markdown and HTML documents
- We can *xtable* class objects through the print command and add arguments to customize the table format 
- `type`  can be `htlm` or `latex` 

``` {r, results='asis', echo=TRUE}
# Load xtable
library(xtable)
# Create an xtable object from M1
M1Table <- xtable(M1, caption = "Linear Regression, Dependent
                      Variable: Exam Score",
                      label = "BasicXtableSummary",
                      digits = 1)
# Create HTML summary table of M1Table
print.xtable(M1Table, type = "html", caption.placement = "top")
```

## Creating Tables from Supported Class R Objects

### `texreg` for LaTeX and HTML

- Creates output tables with estimates from multiple statistical models 
- Supports more model object types than `xtable` 

``` {r, results='asis'}
# Estimated nested regression models
M1 <- lm(Examination ~ Education, data = swiss)
M2 <- lm(Examination ~ Education + Agriculture, data = swiss)
M3 <- lm(Examination ~ Education + Agriculture + Catholic, data = swiss)
M4 <- lm(Examination ~ Education + Agriculture + Catholic + Infant.Mortality, data = swiss)
M5 <- lm(Examination ~ Education + Agriculture + Catholic + Infant.Mortality + Fertility, data = swiss)
```

#### `textreg` for LaTex

```{r, comments='##'}
# Load texreg package
library(texreg)
# Create custom coefficient names
cust_coef <- c('(Intercept)', 'Education', 'Agriculture', 'Catholic', 'Infant Mortality','Fertility')
# Create nested regression model table
texreg(list(M1, M2, M3, M4, M5),
      caption = 'Nested Estimates Table with \\emph{texreg}',
      caption.above = TRUE,
      label = 'Basic_texregTable',
      custom.coef.names = cust_coef)

```

#### `textreg` for HTML

```{r, results='asis'}
library (texreg)
htmlreg(list(M1, M2, M3, M4, M5),
        caption = 'Nested Estimates Table in HTML Document',
        caption.above = TRUE,
        custom.coef.names = cust_coef)
```

## Creating Variable Description documents with xtable

```{r, results='asis'}
# Create variable vector from column names
Variable <- names(swiss)

# Create variable description vector
Description <- c("common standardized fertility measures",
"% of males involved in agriculture as occupation",
"% draftees receiving highest mark on army examination",
"% education beyond primary school for draftees",
"% 'catholic' (as opposed to 'protestant')",
"% live births who live less-than 1 year"
)

# Combine Variable and Description variables into a matrix
DescriptionsBound <- cbind(Variable, Description)

# Create an xtable object from DescriptionsBound
DescriptionsTable <- xtable(DescriptionsBound)

# Format table in HTML
DescriptTable <- print.xtable(DescriptionsTable, type = "html")

# Create variable description file
cat("# Swiss Data Variable Descriptions \n",
    "### Source: Mosteller and Tukey, (1977) \n",
    DescriptTable,
    file = "SwissVariableDescriptions.md"
)
```