---
title: "Chpt_13"
author: "C Degnin"
date: "September 14, 2016"
output: slidy_presentation
---


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, message=FALSE, warning=TRUE)
```
##Formating for Presentations in Markdown:
You can regulated the size of your font in Markdown using 'headings' marks,
but in *HTML presentations* a single hashtage indicates 'new slide':
```
#     A level one heading (not shown)
##    A level two heading
###   A level three heading
####  A level four heading
```

## A level two heading
### A level three heading
#### A level four heading


Similarly, three of more 'equals' (level one heading) or 'dashes' (level two heading) changes your heading in Markdown; 'equals' specifies 'new slide' in *HTML presentations*
```
A level one heading (not shown)
===================

A level two heading 
--------------
```

A level two heading 
--------------



Creating new paragraphs
===============================================

## To create a new paragraph: separate text with a blank line:

```
This is the first paragraph

This is the second paragraph.
This is **still** the second paragraph

This is the third paragraph.   Adding two spaces SHOULD move this to the next line down - but it didn't work.
```
This is the first paragraph

This is the second paragraph.
This is **still** the second paragraph

This is the third paragraph.   Adding two spaces SHOULD move this to the next line down - but it didn'work.



Escape Characters and Lists
======================================================
```
use the backslash (\) as an escape character (to use \*, #, \). 
Although not required in rmarkdown, use &amp for Ampersand (&) and &lt for less-than (<) in HTML. 
```

```
 For lists use one dash or astericks (items)
 
 * Item one
 * Item two
 * Item three
 
 or numbers with periods  (ordered)
 
 1. Item one
 2. Item two
 3. Item three
 
 Note:  The list must be separated from the rest of the code by a blank line.
```
use one dash or astericks 
 
 * Item one
 * Item two
 * Item three
 
 or numbers with periods
 
 1. Item one
 2. Item two
 3. Item three

 
Other symbols and text options
=======================================================
```
*Use the astricks to italize words*.
Use **TWO astricks** to bold words 
(this doesn't work when adjacent to an italicized text)
```

*Use the astricks to italize words*.
Use **TWO astricks** to bold words (this doesn't work adjacent to an italicized text)

## To create **hyperlinks** 

```
use  [LINK_TEX](URL) syntax
For example:
[motion chart](http://127.0.0.1:21393/custom/googleVis/MotionChartID2a6c1ec73d30.html) 
```

use [LINK_TEX](URL) syntax: [motion chart](http://127.0.0.1:18646/custom/googleVis/MotionChartID11cc19815c5e.html) 
(Note this works best in browser mode) 


Overview of googleVis
==============================
[googleVis](https://code.google.com/archive/p/google-motion-charts-with-r/) is a package for R and provides an interface between R and the [Google Chart Tools](https://developers.google.com/chart/)

The functions of the package allow users to visualize data with the Google Chart Tools without uploading their data to Google

The output of googleVis functions is html code that contains the data and references to JavaScript functions hosted by Google

To view the output a browser with an internet connection is required, the actual chart is rendered in the browser; some charts require Flash

See also: Using the Google Visualisation API with R, [The R Journal, 3(2):40-44, December 2011](https://journal.r-project.org/archive/2011-2/RJournal_2011-2_Gesmann+de~Castillo.pdf) and googleVis [package vignette](https://cran.r-project.org/web/packages/googleVis/vignettes/googleVis.pdf)


#Introduction to Google Chart Tools

Google Chart Tools provide a way to visualize data on web sites

The API makes it easy to create interactive charts

It uses JavaScript and DataTable / JSON as input

Output is either HTML5/SVG or Flash

Browser with internet connection required to display chart

You can review the Google [Terms of Service here](https://developers.google.com/terms/)


#Structure of Google Charts
The chart code has five generic parts:

1.	References to Google's AJAX and Visualisation API

2.	Data to visualize as a DataTable

3.	Instance call to create the chart

4.	Method call to draw the chart including options

5.	HTML `<div>` element to add the chart to the page

# How is this done?
.	Transform data into JSON object 
.	Wrap some HTML and JavaScript around it 
.	Thus, googleVis started life in August 2010
.	Motion chart example:

``` {r , echo=TRUE, include = TRUE}
suppressPackageStartupMessages(library(googleVis))
plot(gvisMotionChart(Fruits, "Fruit", "Year", options = list(width = 600, height = 400)))
```

[motion chart](http://127.0.0.1:21393/custom/googleVis/MotionChartID2a6cb394b90.html)

[TED presentation](http://decastillo.github.io/googleVis_Tutorial/#5)




Creating an interactive map - present from the web
============================================

```{r LoadPackages, include=TRUE}
# Load required packages
library(devtools, googleVis)
```

We can use the [googleVis](http://code.google.com/p/google-motion-charts-with-r/) 
package to create interactive JavaScript tables, charts, and maps. Here is an 
example of how to create a map with **googleVis**'s `gvisGeoChart` function. 

First download some data from [GitHub](http://bit.ly/V0ldsf). See chapters 
6 and 7 for details about this data as well as the 
[variable description page](https://github.com/christophergandrud/Rep-Res-Examples/blob/master/DataGather_Merge/MainData_VariableDescriptions.md).

Create geo map of global fertilizer consumption for 2003

The data is loaded from [GitHub](http://bit.ly/V0ldsf)

The data gathering process used to create this data set
is completely reproducible. For more information see:
http://bit.ly/YnMKBG

-----------

## Fertilizer Consumption (kilograms per hectare of arable land) in 2003
### Data from [World Bank](http://data.worldbank.org/indicator/AG.CON.FERT.ZS)


```{r CreategvisGeoChart, echo=FALSE, message=FALSE, results='asis', include = TRUE}

source_url("http://bit.ly/VNnZxS")
```

-----------




 
Between the steps - creating the map
=============================

If you need to install packages inside knitr statement 
**you must include code for cran mirror:**

install.packages("devtools", repos = 'http://cran.us.r-project.org')
 

```{r}
library(devtools)
library(googleVis)
library(repmis)
```


Load data from GitHub (http://bit.ly/V0ldsf)
The data gathering process used to create this data set
is completely reproducible. For more information see:
http://bit.ly/YnMKBG

```{r}
MainData <- source_data("http://bit.ly/V0ldsf")

```

Subset MainData so that it only includes 2003

```{r subdata1}
SubData <- subset(MainData, year == 2003)
```

Remove values of FertilizerConsumption less than 0.1

```{r subdata2}
SubData <- subset(SubData, FertilizerConsumption > 0.1)
```

# Completing the dataset:
Find the natural logarithm of FertilizerConsumption.
Round the results to one decimal digit.
```{r subdata3}
SubData$LogConsumption <- round(log(SubData$FertilizerConsumption),
                                digits = 1)

```



Make a map of Fertilizer Consumption

```{r map }
FCMap <- gvisGeoChart(data = SubData,
                      locationvar = "iso2c",
                      colorvar = "LogConsumption",
                      options = list(
                          colors = "['#ECE7F2', '#A6BDDB', '#2B8CBE']",
                          width = "780px",
                          height = "500px"))

```



# Print FCMap
```{r print, results='asis'}
# Print FCMap
print(FCMap, tag = "chart")
```





 