---
title: "Intro to RStudio"
author: "Jessica Minnier"
date: "August 9, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Rstudio tips

Run a line or run the whole document.

## Keyboard shortcuts

Run a line:
```
ctrl-enter
cmd-enter
```

Search all files in a directory:
```
ctrl-shift-f
cmd-shift-f
```

[Here are some others](https://github.com/SISBID/Module1/blob/gh-pages/labs/rstudio-tour.md)

  * `Ctrl + L` = clear commands
  * `Ctrl + 2` = move cursor to console
  * `Ctrl + 1` = move cursor to script page

## Code autocompletion and help

tab inside parentheses:

```{r,eval=FALSE}

library(ggplot2)
ggplot()
```


## File path autocompletion

```{r, eval=FALSE}
try(source(""))

```

## [Navigating code](https://support.rstudio.com/hc/en-us/articles/200710523-Navigating-Code)

Go to file/function (in toolbar) and search "myprint" or "examplecode" and "add_num"

```{r}
myprint <- function(i) print(i)
```

Navigate with tables of contents within file (see below file, or top right)

Code folding

```{r long_function}
myfun <- function(a,b,c) {
  print(a)
  d = a*b
  print(d)
}
```

## Installing packages is easy

Tools --> install packages


## Foldable regions

You can fold sections of your code within an R script.

```{r}
## fold this -------------------

### or this =======================
```

## Insert sections

code--> insert section (with .R file) cmd-shift-i
code--> insert chunk (with .Rmd file) cmd-alt-i

```{r}

```

## [Code snippets](https://support.rstudio.com/hc/en-us/articles/204463668-Code-Snippets)


press tab after these:

```{r, eval=FALSE}
appl

ts
```

## Addins

```{r addins, eval=FALSE}
install.packages("addinslist")
```

## View Data Sets

```{r, eval=FALSE}
View(mtcars)
```

## Debugging + [Code Diagnostics](https://support.rstudio.com/hc/en-us/articles/205753617-Code-Diagnostics)

Rstudio --> preferences --> Code --> debugging -->
check all boxes

R studio can debug some of your code in real time, also has other debugging tools built in,
such as `browser()`.

[A longer tutorial is here](https://support.rstudio.com/hc/en-us/articles/205612627-Debugging-with-RStudio)

## Start a shell

Tools -> Shell


## New features with every release

Currently in the "preview" release is built in [profiling](https://support.rstudio.com/hc/en-us/articles/218221837-Profiling-with-RStudio) (determining the time each part of code takes to run)

## [Customizing Rstudio](https://support.rstudio.com/hc/en-us/articles/200549016-Customizing-RStudio)

# Reproducibility

- Avoid using mouse to do anything
- Avoid saving workspace or `Rdata` files, just rerun scripts
- Good idea to clear workspace in the beginning of files

```{r}
rm(list=ls())
```

- Use relative file paths
```{r, eval=FALSE}
mydata = read.table("../data/myfile.csv")
```


